---
title: "First day of EDA"
author: "Brock Kamrath"
date: "2023-05-02"
output: html_notebook
---
# Purpose
I need to better understand the data provided to me by NCWQR. Here, I am mainly interested in SRP data. There are several objectives here. First, I need to determine how many days have data in each water year. 

# Method 
Data visualization was conducted using ggplot2


# Results
## Initial data visualization
The SRP data showed an increase in daily loads in the past 5 years.
```{r}
library(tidyverse)
library(ggplot2)
library(here)
library(knitr)

source(here("code","functions","01_data_processing.R"))

df <- get_data(here("data","raw","2023_05_02_Raw-Honey-Wide.csv"))
df_long <- raw_long(df)


# filter the data for SRP and count values per year
count_wy <- df_long %>%
  mutate(wy = water_year(Date)) %>%
  filter(pol == "SRP" & wy >=2000 & wy <= 2021) %>%
  mutate(Water_Year = as.factor(wy)) %>%
  group_by(Water_Year) %>%
    summarise(n = sum(!is.na(flux_kgyr)))


kable(count_wy)

# then summarise the load estimates

df_long %>%
  mutate(wy = water_year(Date)) %>%
filter(pol == "SRP" & wy >=2000 & wy <= 2021)%>%
ggplot(aes(x = as.factor(wy), y = log(flux_kgyr))) + 
  geom_boxplot()+
  stat_summary(fun=mean, geom="line", aes(group=1))  + 
stat_summary(fun=mean, geom="point")+
  theme_bw()

```


```{r}
# calculate geometric mean for each month and look at changes
sum_stat <- df_long %>%
  filter(pol == "SRP") %>%
    mutate(wy = water_year(Date)) %>%
      filter(wy >= 2000 & wy <= 2021) %>%
      mutate(month = month(Date),
             interval = as.numeric(cut(wy,4))) %>%
        group_by(interval,wy,month) %>%
          summarise(gmean = geomMean(flux_kgyr))


sum_stat %>%
  ggplot(aes(x=wy, y = gmean))+
  geom_point(aes(color = interval))+
  geom_smooth(se = FALSE)+
  facet_wrap(~month, scales = "free")


#summaries all values within interval
df_long %>%
  filter(pol == "SRP") %>%
    mutate(wy = water_year(Date)) %>%
      filter(wy >= 2000 & wy <= 2021) %>%
      mutate(month = month(Date),
             interval = as.numeric(cut(wy,4))) %>%
  ggplot() +
  geom_boxplot(aes(x = as.factor(interval), y = log(flux_kgyr)))+
  facet_wrap(~month)
  
```